\section{Executive Summary}

The Christ Embassy Ireland Zone Church Reporting Platform is a unified web-based
solution designed to streamline record-keeping, reporting, and oversight across
key aspects of ministry administration. The platform will consist of four
dedicated sub-domains:

\begin{itemize}
    \item \textbf{Registry} --- for managing member information, first-timer
          registrations, service attendance, and departmental assignments.
    \item \textbf{Finance} --- for logging monetary giving such as offerings,
          tithes, partnerships, and other contributions, including support for
          partnership arms like Rhapsody of Realities, The Healing School, the
          InnerCity Mission for Children, and Loveworld Television Ministry.
    \item \textbf{Cells} --- for the reporting of cell ministry activities,
          including attendance, testimonies, outreach results, and cell offerings.
    \item \textbf{Reports} --- a centralised reporting portal for pastors and
          leaders to view dashboards, generate reports, and monitor ministry
          performance across the organisational hierarchy (regions, zones, groups,
          local churches, and outreaches).
\end{itemize}

The system is designed to replace fragmented manual processes and legacy
spreadsheets with a single, reliable, and secure platform. It will improve the
accuracy of records, enhance transparency in financial reporting, and provide
pastors with timely insights for effective decision-making.

The platform will feature a unified backend and a restrictive role-based access
control model. User accounts will be created top-down, with the Zonal Pastor
able to provision accounts for Group Pastors, who can in turn provision for
Local Church Pastors or Coordinators. Pastors can also create portal-specific
roles such as Finance Officers, Church Administrators, or Cell Leaders. Each user's
permissions are limited to their assigned scope within the church hierarchy.

Key objectives of the platform include:

\begin{itemize}
    \item Establishing a single source of truth for membership and financial data.
    \item Ensuring accountability and auditability of records.
    \item Providing actionable insights through dashboards and KPIs on attendance,
          giving, assimilation, and cell ministry effectiveness.
    \item Reducing administrative overhead and reliance on spreadsheets.
    \item Supporting growth and scalability of ministry operations within the zone.
\end{itemize}

This PRD defines the requirements, scope, and success criteria for the
platform. It is intended as a reference for stakeholders, developers, and
technical teams responsible for its implementation.
