\section{KPIs and Success Metrics}

This section defines the Key Performance Indicators (KPIs) and success metrics
for the Christ Embassy Ireland Zone Church Reporting Platform. These metrics
will guide ongoing performance monitoring and confirm whether the system
achieves its intended operational and technical objectives.

\subsection{Overview}

KPIs are grouped into four primary domains aligned with the platform’s portals:
\textbf{Registry}, \textbf{Finance}, \textbf{Cells}, and \textbf{Reports}.
Additionally, cross-cutting technical KPIs measure data integrity, uptime, and
user engagement.

\subsection{Registry KPIs}

\begin{itemize}
    \item \textbf{KPI-R1: Membership Data Completeness} --- percentage of member
    records with all required demographic and contact fields completed.  
    \emph{Target:} $\geq$ 95\%.

    \item \textbf{KPI-R2: Duplicate Records Rate} --- percentage of duplicate or
    merged records over total members.  
    \emph{Target:} $\leq$ 2\%.

    \item \textbf{KPI-R3: First-Timer Registration Timeliness} --- proportion of
    first-timer entries captured within 24 hours of service attendance.  
    \emph{Target:} $\geq$ 90\%.

    \item \textbf{KPI-R4: First-Timer Assimilation Rate} --- percentage of
    first-timers who transition to full membership within 60 days.  
    \emph{Target:} $\geq$ 70\%.

    \item \textbf{KPI-R5: Attendance Reporting Compliance} --- proportion of
    scheduled services with attendance data recorded and approved.  
    \emph{Target:} 100\%.

    \item \textbf{KPI-R6: Foundation School Completion Rate} --- percentage of
    members who complete foundation school within one year of joining.  
    \emph{Target:} $\geq$ 80\%.

    \item \textbf{KPI-R7: Baptism Tracking Rate} --- proportion of baptized
    members with corresponding date recorded in the system.  
    \emph{Target:} 100\%.
\end{itemize}

\subsection{Finance KPIs}

\begin{itemize}
    \item \textbf{KPI-F1: Data Entry Timeliness} --- percentage of service
    offerings entered within 48 hours of collection.  
    \emph{Target:} $\geq$ 95\%.

    \item \textbf{KPI-F2: Verification Timeliness} --- percentage of batches
    verified and locked within one week of entry.  
    \emph{Target:} $\geq$ 95\%.

    \item \textbf{KPI-F3: Reconciliation Accuracy} --- difference between total
    logged offerings and verified deposits per reporting period.  
    \emph{Target:} $\leq$ 1\% variance.

    \item \textbf{KPI-F4: Partnership Fulfilment Rate} --- proportion of active
    partners meeting or exceeding their pledged amounts.  
    \emph{Target:} $\geq$ 85\%.

    \item \textbf{KPI-F5: Fund Mix Visibility} --- availability of clear fund
    distribution reports (tithes, offerings, partnerships, first fruits).  
    \emph{Target:} 100\% availability.

    \item \textbf{KPI-F6: Cell Offering Reconciliation Rate} --- proportion of
    cell-reported offerings successfully reflected in finance summaries.  
    \emph{Target:} $\geq$ 95\%.
\end{itemize}

\subsection{Cells KPIs}

\begin{itemize}
    \item \textbf{KPI-C1: Reporting Compliance} --- percentage of active cells
    that submit at least one report per week.  
    \emph{Target:} $\geq$ 95\%.

    \item \textbf{KPI-C2: Attendance Accuracy} --- variance between average cell
    attendance and verified church service attendance ratios.  
    \emph{Target:} $\leq$ 10\%.

    \item \textbf{KPI-C3: Outreach Effectiveness} --- number of first-timers and
    new converts reported per 100 active members.  
    \emph{Target:} growth trend $\geq$ +10\% quarter-over-quarter.

    \item \textbf{KPI-C4: Testimony Capture Rate} --- proportion of cell reports
    containing at least one recorded testimony.  
    \emph{Target:} $\geq$ 50\%.

    \item \textbf{KPI-C5: Leader Submission Timeliness} --- percentage of cell
    reports submitted within 24 hours of meeting.  
    \emph{Target:} $\geq$ 90\%.
\end{itemize}

\subsection{Reports \& Oversight KPIs}

\begin{itemize}
    \item \textbf{KPI-P1: Dashboard Availability} --- uptime of the reports
    portal and dashboard services.  
    \emph{Target:} 99.9\%.

    \item \textbf{KPI-P2: Report Generation Speed} --- average generation time
    for standard dashboards and exports.  
    \emph{Target:} $\leq$ 2 seconds for dashboards, $\leq$ 10 seconds for exports.

    \item \textbf{KPI-P3: Hierarchical Accuracy} --- correctness of aggregated
    metrics when drilling down from zone to church level.  
    \emph{Target:} 100\% consistency.

    \item \textbf{KPI-P4: Data Refresh Latency} --- delay between entry or
    update in sub-portals and reflection in dashboards.  
    \emph{Target:} $\leq$ 5 minutes.

    \item \textbf{KPI-P5: Scheduled Report Reliability} --- success rate of
    automated scheduled report deliveries.  
    \emph{Target:} $\geq$ 99\%.
\end{itemize}

\subsection{Cross-Domain Technical KPIs}

\begin{itemize}
    \item \textbf{KPI-T1: System Uptime} --- total platform uptime excluding
    maintenance windows.  
    \emph{Target:} 99.9\%.

    \item \textbf{KPI-T2: API Latency} --- median and 95th percentile latency of
    API endpoints under nominal load.  
    \emph{Target:} P95 $\leq$ 300 ms.

    \item \textbf{KPI-T3: Error Rate} --- percentage of failed API requests
    (5xx or unhandled 4xx) over total requests.  
    \emph{Target:} $\leq$ 0.5\%.

    \item \textbf{KPI-T4: Data Integrity Score} --- proportion of records
    passing consistency checks and referential integrity validation.  
    \emph{Target:} $\geq$ 99.5\%.

    \item \textbf{KPI-T5: Backup Validation Success} --- successful restore
    verification rate per quarter.  
    \emph{Target:} 100\%.

    \item \textbf{KPI-T6: Audit Log Coverage} --- proportion of sensitive
    actions successfully captured in audit logs.  
    \emph{Target:} 100\%.
\end{itemize}

\subsection{Measurement \& Monitoring}

\begin{itemize}
    \item KPI metrics shall be automatically collected and displayed in the
    administrative reporting dashboard for zone leadership and technical staff.
    \item Historical KPI data shall be stored for at least 12 months for trend
    analysis.
    \item Alerts shall be configured for threshold breaches (e.g., missing
    service reports, delayed batch verification).
    \item KPI data shall also be exportable in CSV format for external analysis.
\end{itemize}
