\section{Objectives and Goals}

The primary objective of the Christ Embassy Ireland Zone Church Management Platform is
to provide a unified, secure, and scalable backend for managing people,
finance, cells, and reports across the zone. The system is designed to replace
manual processes and fragmented spreadsheets with structured data entry,
controlled access, and automated reporting.

\subsection*{Goals}
\begin{itemize}
    \item Establish a single source of truth for membership, attendance, financial
          contributions, and cell ministry reports.
    \item Ensure data integrity and accountability through role-based access control and
          audit logging.
    \item Provide timely and accurate dashboards and KPIs for pastors at different levels
          of oversight (zone, group, church, cell).
    \item Standardise data entry formats to reduce duplication and errors.
    \item Enable efficient migration of legacy records into a unified database.
    \item Support growth through modular design and scalability.
\end{itemize}

\subsection*{Non-Goals}
\begin{itemize}
    \item Payment processing --- the platform records giving, but does not process
          transactions directly.
    \item Event scheduling, media management, or volunteer rostering --- these are out of
          scope for the MVP.
    \item Replacing KingsPay or other existing giving platforms --- the focus is on
          record-keeping and reporting.
    \item Advanced CRM or mass communication features --- only essential follow-up fields
          and contact flags will be included.
\end{itemize}
