\section{Risks and Assumptions}

This section identifies the major risks, dependencies, and underlying
assumptions associated with the Christ Embassy Ireland Zone Church Reporting
Platform. Each risk includes an impact assessment and mitigation strategy.

\subsection{Project Assumptions}

\begin{enumerate}
    \item The system will initially be deployed for the \textbf{Ireland Zone}
    and may later be replicated for other zones.
    \item All participating churches, groups, and cells will maintain stable
    internet access sufficient for web-based data entry.
    \item Existing spreadsheets and records are sufficiently structured to
    support automated import after field mapping and validation.
    \item The zone has designated personnel for each portal:
    Church Administrators, Finance Officers, Cell Leaders, and Report Viewers.
    \item All users will undergo basic onboarding and training before launch.
    \item Zone leadership will define approved organisational hierarchies and
    naming conventions prior to data migration.
    \item Financial records are entered manually for reporting; no real payment
    processing is performed on the platform.
    \item Access control will follow the hierarchical structure
    (Zonal → Group → Church → Cell) without exceptions.
    \item A central technical administrator or IT officer will manage backups,
    infrastructure, and access provisioning at the zone level.
\end{enumerate}

\subsection{Project Risks and Mitigation Strategies}

\paragraph{1. Data Quality Risk}
\begin{itemize}
    \item \textbf{Description:} Legacy spreadsheets may contain inconsistent or
    incomplete data.
    \item \textbf{Impact:} Duplicate or invalid records could distort reporting
    accuracy.
    \item \textbf{Mitigation:} Implement pre-import validation, deduplication
    tools, and review dashboards for anomalies.
\end{itemize}

\paragraph{2. User Adoption Risk}
\begin{itemize}
    \item \textbf{Description:} Transitioning from manual processes to a digital
    system may face initial resistance from some users.
    \item \textbf{Impact:} Incomplete reporting or low data entry compliance.
    \item \textbf{Mitigation:} Provide short training sessions, in-portal help,
    and simplified mobile access for clerks and cell leaders.
\end{itemize}

\paragraph{3. Connectivity Risk}
\begin{itemize}
    \item \textbf{Description:} Some outreaches or churches may experience poor
    network connectivity.
    \item \textbf{Impact:} Delayed data entry or missed attendance submissions.
    \item \textbf{Mitigation:} Support offline entry templates or lightweight
    mobile interfaces that sync when connectivity is restored.
\end{itemize}

\paragraph{4. Access Control Misconfiguration}
\begin{itemize}
    \item \textbf{Description:} Incorrect assignment of scopes or roles may
    expose data beyond intended boundaries.
    \item \textbf{Impact:} Breach of data privacy or loss of trust in system
    security.
    \item \textbf{Mitigation:} Use guided user provisioning wizards, predefined
    role templates, and periodic permission audits.
\end{itemize}

\paragraph{5. Financial Data Entry Errors}
\begin{itemize}
    \item \textbf{Description:} Manual entry of offerings or partnerships may
    introduce typos or misallocations.
    \item \textbf{Impact:} Reporting inaccuracies or reconciliation variance.
    \item \textbf{Mitigation:} Dual control for batch locking, field-level
    validation (numeric ranges, required fund), and audit trails.
\end{itemize}

\paragraph{6. Delayed Report Generation}
\begin{itemize}
    \item \textbf{Description:} Large reports may take long to generate during
    peak hours.
    \item \textbf{Impact:} Poor user experience and delayed insights for pastors.
    \item \textbf{Mitigation:} Use pre-aggregated summary tables, background job
    queues, and read replicas for analytics.
\end{itemize}

\paragraph{7. Data Loss or Corruption}
\begin{itemize}
    \item \textbf{Description:} Accidental deletion, database error, or hardware
    failure could compromise stored data.
    \item \textbf{Impact:} High; loss of financial or membership records.
    \item \textbf{Mitigation:} Daily encrypted backups, multi-region storage,
    and quarterly restore tests.
\end{itemize}

\paragraph{8. Audit Log Overflow or Mismanagement}
\begin{itemize}
    \item \textbf{Description:} Unbounded audit log growth may degrade
    performance or consume storage.
    \item \textbf{Impact:} Reduced database efficiency.
    \item \textbf{Mitigation:} Partition audit tables by date, archive older
    records to cold storage after 12 months.
\end{itemize}

\paragraph{9. Version Drift Between Portals}
\begin{itemize}
    \item \textbf{Description:} If portal frontends evolve asynchronously, UI
    inconsistencies or incompatible API calls may occur.
    \item \textbf{Impact:} Broken functionality or reporting errors.
    \item \textbf{Mitigation:} Maintain shared component libraries and strict
    version tagging across all sub-domains.
\end{itemize}

\paragraph{10. Leadership Reporting Delays}
\begin{itemize}
    \item \textbf{Description:} Pastors may not receive timely insights if
    lower-level data entry is delayed.
    \item \textbf{Impact:} Leadership decisions based on incomplete data.
    \item \textbf{Mitigation:} Scheduled reporting reminders, compliance KPIs,
    and automated notifications for missing submissions.
\end{itemize}

\subsection{Technical Risks}

\paragraph{1. Scalability Constraints}
\begin{itemize}
    \item \textbf{Description:} Rapid growth in churches or records may outpace
    single-database performance.
    \item \textbf{Mitigation:} Deploy read replicas for reporting, introduce
    partitioning on date/org\_unit\_id, and plan migration path to distributed
    analytics store if necessary.
\end{itemize}

\paragraph{2. Background Job Failures}
\begin{itemize}
    \item \textbf{Description:} Import, export, or scheduled jobs may fail
    silently due to worker errors.
    \item \textbf{Mitigation:} Implement job retry policies, error queues, and
    alerting dashboards for failed tasks.
\end{itemize}

\paragraph{3. API Performance Degradation}
\begin{itemize}
    \item \textbf{Description:} Inefficient queries or unindexed filters may
    degrade API latency under load.
    \item \textbf{Mitigation:} Continuous monitoring of query plans and use of
    cached summary tables for heavy endpoints.
\end{itemize}

\subsection{Operational Risks}

\paragraph{1. Staff Turnover}
\begin{itemize}
    \item \textbf{Description:} Departure of trained users may reduce data
    entry consistency.
    \item \textbf{Mitigation:} Maintain user manuals, assign backup clerks, and
    provide refresher sessions.
\end{itemize}

\paragraph{2. Misinterpretation of Reports}
\begin{itemize}
    \item \textbf{Description:} Pastors or finance teams may misread dashboard
    data if untrained in metric definitions.
    \item \textbf{Mitigation:} Include tooltips and embedded data definitions on
    dashboards; provide glossary in appendix.
\end{itemize}

\paragraph{3. Governance Risk}
\begin{itemize}
    \item \textbf{Description:} Absence of clear data ownership could lead to
    disputes or inconsistent updates.
    \item \textbf{Mitigation:} Define roles and responsibilities clearly within
    each church and group; enforce audit-based accountability.
\end{itemize}

\subsection{Residual Risk Statement}

While technical measures and governance processes significantly reduce
likelihood and impact, residual risk remains in user compliance and operational
discipline. Continuous monitoring, periodic training, and consistent leadership
oversight will be essential to maintain data quality and trust in the system.
