\section{Project Scope}

The Christ Embassy Zone Church Management Platform will deliver a unified
backend with four dedicated sub-domains, each optimised for a specific
operational area. All modules share a common data model and a single access
control system.

\subsection*{In Scope}

\begin{itemize}
    \item \textbf{Registry Portal} (\texttt{registry.<zone>.ce.church})
          \begin{itemize}
              \item Member information management (personal details, contact info, marital status,
                    foundation school completion, baptism date).
              \item First-timer registration and follow-up status tracking.
              \item Service attendance records (men, women, teens, kids, first timers, new
                    converts).
              \item Departmental assignments and leadership roles.
          \end{itemize}

    \item \textbf{Finance Portal} (\texttt{finance.<zone>.ce.church})
          \begin{itemize}
              \item Recording of offerings, tithes, seeds, first-fruits, and other giving types.
              \item Partnership arm support (Rhapsody of Realities, Healing School, InnerCity
                    Mission, Loveworld TV, plus custom campaigns).
              \item Linking transactions to members, cells, or external givers.
              \item Batch locking and verification for service-linked entries.
              \item Simple reconciliation and reporting views.
          \end{itemize}

    \item \textbf{Cells Portal} (\texttt{cells.<zone>.ce.church})
          \begin{itemize}
              \item Cell leader reporting (attendance, first timers, new converts, testimonies,
                    offerings, meeting type).
              \item Linking members to cells and designating leadership.
              \item Escalation flags for pastoral follow-up.
          \end{itemize}

    \item \textbf{Reports Portal} (\texttt{reports.<zone>.ce.church})
          \begin{itemize}
              \item Dashboards and KPIs covering membership, finance, cells, and attendance.
              \item Drill-down and roll-up across the organisational hierarchy (region, zone,
                    group, church, cell).
              \item CSV, Excel, and PDF exports of standard reports.
              \item Scheduled email reports.
          \end{itemize}
\end{itemize}

\subsection*{Out of Scope}

\begin{itemize}
    \item Direct payment processing --- giving is recorded, not transacted.
    \item Event scheduling, volunteer management, and media handling.
    \item Advanced CRM features (mass messaging, marketing automation).
    \item Integration with external accounting software (exports only in MVP).
    \item Mobile native applications --- all portals will be responsive web apps.
\end{itemize}
